Background Music Will not Significantly Distract Attention

### 2021-04-25 Output/paper/paper.Rmd
update

---
title: "Background Music Will not Distract Students' Attention"
author: "Hong Pan"
date: "`r format(Sys.time(), '%B %d %Y')`"
output: 
  bookdown::pdf_document2:
    toc: yes

abstract: "Music has gradually become an essential part of our lives, and it has become a common trend for most students to choose to listen to music while studying. This paper examines whether the background music affects students' attention level by conducting a within-subjects experiment to compare the time of finishing the Trail Making Test with and without music. From the liner mixed mode, the result shows that background music does not significantly impact attention level. Apart from that, male students tend to be more easily distracted by the background music. \\par

 \\textbf{Keywords:} background music, attention, trail making test, executive functioning skills"

@@

Trail Making Test, which is used to measure executive functioning skills, was utilized to study whether background music would influence students’ attention level. Executive functions refer to the mental processes needed when people have to concentrate or pay attention, which is "a specific set of attention-regulation skills involved in conscious goal-directed problem-solving.[@executive]” Defectiveness of executive functioning skills will make students have a hard time paying attention[@Understood]. In the Trail Making Test, each round consists of several dots numbered from 1 to 12 and lettered A to J, randomly distributed on a page. The participants are required to connect the circles as soon as possible.^[The order should alternate between numbers and letters, i.e. 1, A, 2, B,...]

The within-subjects experiment is conducted by randomly selected students from one specific school, and each participant needs to complete two rounds of tests, with and without music.  The dataset created by Arriken Worsley[@dataset] records the time of accomplishing the test, gender of the student, music favourability, and the order of test which is starting with the music or without the music. The descriptive analysis results show that students performed better without music. In addition, whatever the order of the test, the second round will be faster on average. Furthermore, background music has the biggest influence on male's executive functioning skills. Because the dataset is longitudinal, the linear mixed model is applied to do the statistical analysis. The result from the model illustrates that the students’ attention level is not significantly impacted by listening to the music. The paper concludes that the background music does not significantly impact students' attention level. In this study, the practice effect^[People improve their scores simply by taking the same test multiple times. (Kendra Cherry,2019)]  may exist since the students need to do the test twice continuously. Finally, the interactive graphs display on Shinyapp^[The Shinyapp link for the interactive graphs is https://hongtrista.shinyapps.io/Shinyapp/?_ga=2.86381278.759236858.1619203106-1272735195.1619203106 ] allow the audience to explore the data distribution by gender, test type and rating.

The rest of this article is structured as follows. In the data section, the study design, including the experiment process, sampling method, and the dataset variables, will be explained. In the model section, the model selection will be discussed, which aims to investigate how background music impacts students’ attention level. In the result section, the descriptive and statistical analysis detail will be presented. Finally, in the discussion section, the main findings, limitations and direction of future research will be discussed. In the Appendix, the screenshot for the survey and the real Trail Making Test used for the experiment will be shown.

@@

### Intervention
The purpose of the experiment is to examine whether background music affects the attention of students. The design of the experiment was to randomly select some students from the school to complete the "Trail Making Test" with and without background music and record the time of completion. For this experiment, the intervention is the presence of background music. Experiment designer Arriken Worsley wanted to determine whether background music would affect students' attention level by studying whether students took longer to complete the test with background music than without background music. If the result is that the time for the student to complete the test with background music is longer than the time without background music, it means that the background music affects the student's attention to a certain extent. Otherwise, background music will not affect students' attention. The experiment can help the target audience develop some basic understanding of whether music impacts attention and correctly decide whether to listen to music in future studies or avoid inefficiency.

### Procedure
As the paper aims to understand whether background music would affect students' attention level, all students in the world would be considered as the target population. The frame population is all students from the school. 47 students being randomly selected from the school as a sampled population to finish the experiment. The detailed procedure will be explained as followed.

The 47 students chosen randomly from the school will be randomly separated into two groups to start the Trail Making Test in a different order. 27 students started the experiment in a quiet environment, and the remaining 20 students started with the existence of background music. Figure \@ref(fig:test1) and Figure \@ref(fig:test2) in the appendix show two different Trail Making Test used in different test environments during the experiment. Each test contains 23 randomly distributed dots, including 12 numeric points from 1 to 12 and 11 letter points from A to K. Participants must start connecting from 1 and connect these points to the final point in alternating ascending order, such as 1, A, 2, B, 3, C, et cetera. During the process, the line should not cross over. The background music, the intervention,  while playing during the experiment will be the song Panda by Designer because, during the experiment period, it was at the top of the billboard and was the most popular music.

@@

The paper[@female] written by Bruce Goldman illustrates that females are good at concentration and not easily being distracted compared with males. In order to study whether background music will affect students' attention levels differently due to different genders, Figure \@ref(fig:gendertime) is drawn. Overall, the conclusion from the graph goes along with the conclusion Bruce Goldman made[@female]: females are less affected by the intervention. At the same time, males' attention levels are easily affected. The descriptive analysis result shows that background music increased the average time for female and transexual students to complete the test by about one second. In contrast, background music influences male students' average time to complete the test increased by 5 seconds. Moreover, whether with or without background music, the average completion time for men is the longest, while the average time for transexuals is the shortest.
```{r gendertime, fig.cap="Average time for different genders to complete the test by test type",echo=FALSE,fig.height=4.5,fig.width=5,message=FALSE}
d1 <- dataset %>% 
        group_by(Gender,Test_Type) %>% 
        summarise(mean_time = mean(Time_Sec))

@@

```{r r3, fig.cap="Comparing Model by ANOVA",echo=FALSE,warning=FALSE,message=FALSE,comment=NA}
anova(m1,m2,m3,m4,m5)%>%
  kbl(caption = 'To Find Out the Best Final Model by ANOVA',
      align = "cccc") %>%
  kable_styling(latex_options = "hold_position")
```

## Model Assessment Result
Root Mean Square Error(RMSE) is used to evaluate the quality of the model. RMSE stands for the standard deviation of the residuals and it helps to tell how concentrated the data is around the best fit line[@RMSE]. In Table \@ref(tab:r4), the RMSE value of the final model, model 2, is 0.135 which a good sign for the model and illustrates that the model has a good ability to predict the data accurately. The R-square is applied to assess the model as well. R-square is a statistical measure of how close the data are to the fitted regression line[@R-square]. R-square consisted two types, which are marginal R-square and conditional R-square for the linear mixed model. The paper will mainly focused on marginal R-square values, which provides the variance explained by only fixed effect. The marginal R-square for the final model, model 2, shown in Table \@ref(tab:r4) is 0.169, which means the final linear mixed model can explain 16.9% of the variation of the dependent variable.

@@

tibble(
  'RMSE' = 0.134,
  'Marginal R-square' = 0.169,
  'Conditional R-square' = 0.410
) %>%
  kbl(caption = 'Assessment Result',
      align = "cccc") %>%
  kable_styling(latex_options = "hold_position")
```

@@

## Ethics Statement
The experiment conducted in this paper is about learning whether background music will influence students' attention level. The intervention in the study is background music. There is no potential to harm or endanger any participant. Moreover, the designer did not collect any sensitive data. Therefore, no formal ethics approval is required.

@@

### Each Student’s Initial Attention Level is Different
When checking linear mixed model validation, the resulting plot shows that there exist random intercepts and random slopes, which illustrates that each students' initial attention level is different and fixed effects will vary for each student. In other words, the linear mixed model is suitable for this data and can incorporate student-to-student variability in finishing time and improve the ability to describe how fixed effects relate to outcomes.

### Background Music Does not Significantly Affect the Level of Attention of Students
Although in the descriptive analysis, it is found that in the case of background music, the average time for students to complete the test is longer than that without background music. But in the process of comparing the five models, the results show that test_type is not significant among all the models. This shows that it fails to prove that background music will affect the level of attention of students in a statistical sense.

### The Order of Testing will Affect the Time to Complete the Test
When comparing models, the results show that the order significantly affects the time it takes for students to complete the test. In the final best-fit model, the coefficient of order is -10.65, and the 90% confidence interval is from -14.85 to -6.45. The coefficient, -10.65, means, on average, the time for completing the test in second-order is 10.65 seconds faster than the time in the first order when holding other variables constant.

@@

## Weaknesses and Next Step
### Weaknesses

@@

### Next Step
**Between-subject Design with different types of music:**
After expanding the sample size, change the experiment design to a between-subjects study experiment.  Between-subjects design means each participant will only test each situation once, which prevents practice effect by eliminating the transfer of knowledge and shorten testing sessions by allowing each participant only need to do the test once.

@@

>>>>

---
title: "Background Music Will not Significantly Distract Attention"
author: "Hong Pan"
date: "`r format(Sys.time(), '%B %d %Y')`"
output: 
  bookdown::pdf_document2:
    toc: yes

abstract: "Music has gradually become an essential part of our lives, and it has become a common trend for most students to choose to listen to music while studying. This paper used the dataset collected by Arriken Worsley[@dataset] to examine whether the background music affects students' attention level by conducting a within-subjects experiment to compare the time of finishing the Trail Making Test with and without music. The linear mixed model results show that background music does not significantly impact attention level but has a significant negative relationship with the order of the test being done.  \\par

 \\textbf{Keywords:} Background Music, Attention, Trail Making Test, Executive Functioning Skills"

@@

Trail Making Test, which is used to measure executive functioning skills, was utilized to study whether background music would influence students’ attention level. Executive functions refer to the mental processes needed when people have to concentrate or pay attention, which is "a specific set of attention-regulation skills involved in conscious goal-directed problem-solving.[@executive]” Defectiveness of executive functioning skills will make students have a hard time paying attention[@Understood]. In the Trail Making Test, each round consists of several dots numbered from 1 to 12 and lettered A to J, randomly distributed on a page. The participants are required to connect the circles as soon as possible.^[The order should alternate between numbers and letters, i.e. 1, A, 2, B,...] The paper written by Arielle S. Dolegui[@paper1] used repeated measures to design experiments to study the impact of different types of music and playing volume levels on students’ cognitive ability to complete assignments. The result illustrates that students perform the best in silence. Meanwhile, the paper published by Janina and Tina investigates[@paper2] how background music influences learning by conducting a between-subjects design experiment.  The result for their paper demonstrates that background music did no influence performance for learners. Is it true that background music does not affect the ability of students to process information?

The within-subjects experiment is conducted by randomly selected 47 students from Elizabeth Blackburn School of Science, and each participant needs to complete two rounds of tests, with and without music. In this study, the practice effect^[People improve their scores simply by taking the same test multiple times. (Kendra Cherry,2019)]  may exist since the students need to do the test twice continuously. In order to reduce the impact from practice effect, participants were randomly selected to two different groups with different starting order. One group started with a test with background music, while the other group started the test in silence. The dataset created by Arriken Worsley[@dataset] records the time of accomplishing the test, gender of the student, music favourability, and the order of test which is starting with the music or without the music. The descriptive analysis results show that students performed better without music. In addition, whatever the order of the test, the second round will be faster on average. Furthermore, background music has the biggest influence on male's executive functioning skills. Because the dataset is longitudinal, the linear mixed model is applied to do the statistical analysis. The result from the model illustrates that the students’ attention level is not significantly impacted by listening to the music. The paper concludes that the background music does not significantly impact students' attention level.

The rest of this paper is structured as follows. In the data section, the experimental process, sampling method and data set variables will be explained. The descriptive analysis detail will also be presented. In the model section, the reason for choosing the model, model assumptions, and model selection procedure will be discussed. In the result section, the statistical analysis detail and model validation result will be illustrated and explained. In the discussion section, the main findings, limitations, ethics statements, and further research will be discussed.  In the Appendix, the screenshot for the survey and the real Trail Making Test used for the experiment will be shown. Finally, the interactive graphs display on Shinyapp^[The Shinyapp link for the interactive graphs is https://hongtrista.shinyapps.io/Shinyapp/?_ga=2.86381278.759236858.1619203106-1272735195.1619203106 ] allow the audience to explore the data distribution by gender, test type and rating.

@@

### Intervention
The purpose of the experiment is to examine whether background music affects the attention of students. The design of the experiment is to randomly select some students from the school to complete the "Trail Making Test" with and without background music and record the time of completion. For this experiment, the intervention is the presence of background music. Experiment designer Arriken Worsley wants to determine whether background music will affect students' attention level by studying whether students take longer to complete the test with background music than without background music. If the result is that the time for the student to complete the test with background music is longer than the time without background music, it means that the background music affects the student's attention to a certain extent. Otherwise, background music will not affect students' attention. The experiment can help the target audience develop some basic understanding of whether music impacts attention and correctly decide whether to listen to music in future studies or avoid inefficiency.

### Procedure
As the paper aims to understand whether background music will affect students' attention level, all students in the world would be considered as the target population. The frame population is all students from the school. 47 students being randomly selected from the school as a sampled population to finish the experiment. The detailed procedure will be explained as followed.

The 47 students chosen randomly from the school were randomly divided into two groups to start the Trail Making Test in a different order. 27 students started the experiment in a quiet environment, and the remaining 20 students started with the existence of background music. Figure \@ref(fig:test1) and Figure \@ref(fig:test2) in the appendix show two different Trail Making Test used in different test environments during the experiment. Each test contained 23 randomly distributed dots, including 12 numeric points from 1 to 12 and 11 letter points from A to K. Participants must start connecting from 1 and connected these points to the final point in alternating ascending order, such as 1, A, 2, B, 3, C, et cetera. During the process, the line should not cross over. The background music, the intervention,  played during the experiment would be Designer's song, Panda, because, during the experiment period, it was at the top of the billboard and was the most popular music.

@@

The paper[@female] written by Bruce Goldman illustrates that females are good at concentration and not easily being distracted compared with males. In order to study whether background music will affect students' attention levels differently due to different genders, Figure \@ref(fig:gendertime) is drawn. Overall, the conclusion from the graph goes along with the conclusion Bruce Goldman made[@female]: females are less affected by the intervention. At the same time, males' attention levels are easily affected. The descriptive analysis result shows that background music increased the average time for female and transexual students to complete the test by about one second. In contrast, background music influences male students' average time to complete the test increased by 5 seconds. Moreover, whether with or without background music, the average completion time for men is the longest, while the average time for transexuals is the shortest.

```{r gendertime, fig.cap="Average time for different genders to complete the test by test type",echo=FALSE,fig.height=4.5,fig.width=5,message=FALSE}
d1 <- dataset %>% 
        group_by(Gender,Test_Type) %>% 
        summarise(mean_time = mean(Time_Sec))

@@

```{r r3, fig.cap="Comparing Model by ANOVA",echo=FALSE,warning=FALSE,message=FALSE,comment=NA}
round(anova(m1,m2,m3,m4,m5),2)%>%
  kbl(caption = 'To Find Out the Best Final Model by ANOVA',
      align = "cccc") %>%
  kable_styling(latex_options = "hold_position")
```

## Model Assessment Result
Root Mean Square Error(RMSE) is used to evaluate the quality of the model. RMSE stands for the standard deviation of the residuals and it helps to tell how concentrated the data is around the best fit line[@RMSE]. In Table \@ref(tab:r4), the RMSE value of the final model, model 2, is 0.135 which a good sign for the model and illustrates that the model has a good ability to predict the data accurately. The R-square is applied to assess the model as well. R-square is a statistical measure of how close the data are to the fitted regression line[@R-square]. R-square consisted two types, which are marginal R-square and conditional R-square for the linear mixed model. The paper will mainly focused on marginal R-square values, which provides the variance explained by only fixed effect. The marginal R-square for the final model, model 2, shown in Table \@ref(tab:r4) is 0.169, which means the final linear mixed model can explain 16.9% of the variation of the dependent variable. From the statistical result of the final model shown in appendix , the random effect result illustrates that the variance of Id is 43.08 and the variance of residual is 105.41. The variability due to the Id is 29%, which is means 29% within treatment variance is explained by individual differences. It is suitable to use a linear mixed model for this data.

@@

tibble(
  'RMSE' = 0.134,
  'Marginal R-square' = 0.169,
  'Conditional R-square' = 0.410,
  "Variance of Id"=0.29
) %>%
  kbl(caption = 'Assessment Result',
      align = "cccc") %>%
  kable_styling(latex_options = "hold_position")
```

@@

## Ethics statement
The experiment conducted in this paper is about learning whether background music will influence students' attention level. The intervention in the study is background music. There is no potential to harm or endanger any participant. Moreover, the designer did not collect any sensitive data. Therefore, no formal ethics approval is required.

@@

### Each student’s initial attention level is different
When checking linear mixed model validation, the resulting plot shows that there exist random intercepts and random slopes, which illustrates that each students' initial attention level is different and fixed effects will vary for each student. In other words, the linear mixed model is suitable for this data and can incorporate student-to-student variability in finishing time and improve the ability to describe how fixed effects relate to outcomes.

### Background music does not significantly affect the level of attention of students
Although in the descriptive analysis, it is found that in the case of background music, the average time for students to complete the test is longer than that without background music. But in the process of comparing the five models, the results show that test_type is not significant among all the models. This shows that it fails to prove that background music will affect the level of attention of students in a statistical sense.

### The order of testing will affect the time to complete the test
When comparing models, the results show that the order significantly affects the time it takes for students to complete the test. In the final best-fit model, the coefficient of order is -10.65, and the 90% confidence interval is from -14.85 to -6.45. The coefficient, -10.65, means, on average, the time for completing the test in second-order is 10.65 seconds faster than the time in the first order when holding other variables constant.

@@

## Weaknesses and Next Steps
### Weaknesses

@@

### Next Steps
**Between-subject Design with different types of music:**
After expanding the sample size, change the experiment design to a between-subjects study experiment.  Between-subjects design means each participant will only test each situation once, which prevents practice effect by eliminating the transfer of knowledge and shorten testing sessions by allowing each participant only need to do the test once.

@@

## Final Model Output
```{r finalmodel, echo=FALSE, fig.align='center', fig.cap="Final model result",comment=NA}
summary(m2)
```
